\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,bm}
\usepackage[margin=1in]{geometry}

\begin{document}

\section*{Inspection tools for GNN2}

Let $\mathcal{N}$ be the set of bus--phase nodes ($i\in\mathcal{N}$), and let
each sample (snapshot) be indexed by $s$ with time index $t$.

\subsection*{1.\ Spatial inspection (vmag-only GNN)}

Dataset: \texttt{datasets\_gnn2/original}. Model: checkpoint
\texttt{models\_gnn2/original/pf\_identity\_gnn\_vmag\_only\_best.pt}.

For each test snapshot $(s,t)$ and node $i$:
\begin{align*}
  x_i^{(s,t)} &= \bigl(P_{\mathrm{load},i}^{(s,t)},\;
                      Q_{\mathrm{load},i}^{(s,t)},\;
                      P_{\mathrm{pv},i}^{(s,t)}\bigr),\\
  \hat v_i^{(s,t)} &= f_\theta\bigl(x^{(s,t)},\mathcal{G}\bigr)_i,\\
  v_i^{(s,t)} &= |V_i^{(s,t)}|\quad(\texttt{vmag\_pu}),\\
  e_i^{(s,t)} &= \bigl|\hat v_i^{(s,t)}-v_i^{(s,t)}\bigr|.
\end{align*}
The script:
\begin{itemize}
  \item rebuilds the exact train/test split over samples (same seed and test fraction as training),
  \item evaluates $f_\theta$ on all test snapshots,
  \item ranks nodes by $e_i^{(s,t)}$ (worst bus--phase), buses by
        $\max_{ph} e_{(b,ph)}^{(s,t)}$, and snapshots by
        $\frac{1}{|\mathcal{N}|}\sum_i e_i^{(s,t)}$,
  \item produces histograms of $\{e_i^{(s,t)}\}$ and error vs.\ time-of-day
        using \texttt{gnn\_sample\_meta.csv},
  \item writes a detailed table
        $\{i,s,t,\hat v_i^{(s,t)},v_i^{(s,t)},e_i^{(s,t)}\}$
        to \texttt{test\_node\_errors\_detailed.csv}.
\end{itemize}
Thus spatial inspection quantifies and localizes the vmag error of the trained
GNN over the original dataset.

\subsection*{2.\ Dataset/graph inspection}

Dataset directories:
\[
\texttt{datasets\_gnn2/original},\quad
\texttt{datasets\_gnn2/injection},\quad
\texttt{datasets\_gnn2/loadtype},\quad
\texttt{datasets\_gnn2/deltav}.
\]
Common inputs per dataset:
\[
\texttt{gnn\_edges\_phase\_static.csv},\quad
\texttt{gnn\_node\_features\_and\_targets.csv},\quad
\texttt{gnn\_sample\_meta.csv}.
\]

Let $\mathcal{G}=(\mathcal{N},\mathcal{E})$ be the static phase graph with
directed edges $e=(i\to j,ph)$.

\paragraph{(a) Bidirectionality.}
The script checks that
\[
(i\to j,ph)\in\mathcal{E}\ \Longrightarrow\ (j\to i,ph)\in\mathcal{E}
\]
using \texttt{u\_idx}, \texttt{v\_idx}, \texttt{phase} from the edge CSV.

\paragraph{(b) Transformer/regulator edges.}
For each phase $ph\in\{1,2,3\}$ the following pairs must be present in both
directions:
\[
832.ph \leftrightarrow 888.ph,\quad
814.ph \leftrightarrow 814\mathrm{r}.ph,\quad
852.ph \leftrightarrow 852\mathrm{r}.ph.
\]
These correspond to \texttt{XFM1}, \texttt{regA1/2/3}, and
\texttt{regB1/2/3} in the DSS model.

\paragraph{(c) First-sample node features.}
For the smallest sample index $s_0$, the script prints all node rows
$\{x_i^{(s_0,t)},y_i^{(s_0,t)}\}$ from
\texttt{gnn\_node\_features\_and\_targets.csv} to visually confirm indexing and
feature layout.

\paragraph{(d) Aggregated P/Q checks.}
\emph{Injection dataset (Dataset~2).}
For one sample $(s,t)$, with node injections
$(P_{\mathrm{inj},i},Q_{\mathrm{inj},i})$ and upstream grid power
$(P_{\mathrm{grid}},Q_{\mathrm{grid}})$ from sample meta:
\[
P_{\mathrm{res}}=\sum_i P_{\mathrm{inj},i}+P_{\mathrm{grid}},\qquad
Q_{\mathrm{res}}=\sum_i Q_{\mathrm{inj},i}+Q_{\mathrm{grid}},
\]
which should be small and correspond to losses.

\emph{Load-type and delta-V datasets (Datasets 3 and 4).}
Using post-solve load and PV powers and capacitor Q:
\begin{align*}
P_{\mathrm{sys}}(t)
  &=\sum_i P_{\mathrm{load},i}(t)-\sum_i P_{\mathrm{pv},i}(t),\\
Q_{\mathrm{sys}}(t)
  &=\sum_i Q_{\mathrm{load},i}(t)-\sum_i Q_{\mathrm{pv},i}(t)
    -\sum_b Q_{\mathrm{cap}}(b),
\end{align*}
the script recomputes these from node features and compares with the stored
\texttt{p\_sys\_balance\_kw}, \texttt{q\_sys\_balance\_kvar}. For Dataset~4 it
also summarizes the zero-PV and delta-V targets:
\[
|V_i^{(0)}(t)|=\texttt{vmag\_zero\_pv\_pu},\qquad
\Delta|V_i(t)|=\texttt{vmag\_delta\_pu}.
\]

\paragraph{(e) Edge static features.}
Finally, the script prints a sample of rows
$(\texttt{from\_node},\texttt{to\_node},R_{\mathrm{full}},X_{\mathrm{full}},C_{\mathrm{full}})$
so that line vs.\ transformer/regulator impedances (in~ohms) can be checked
directly against the DSS model.

\end{document}

